---
title: "Distributional regression analysis of learning analytics and educational data"
subtitle: Data preparation
output:
  pdf_document: default
  html_notebook: default
---

# Load required libraries
```{r}
# dataset
devtools::install_github("jakubkuzilek/oulad", force = TRUE)
library(oulad) # 

# pipes
library(magrittr)

# data manipulation
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)

# data storing
library(readr)
```

# Selecting data
We focused on the course "FFF" and its presentation "2013J": 
```{r}
module <- "FFF"
presentation <- "2013J"

# information about the course
selected_course <-
  course %>% 
  filter(code_module == module, code_presentation == presentation)

```

The students who did not finished the course (actively withdrawn), were filtered out:
```{r}
# list of students
selected_students <-
  student %>% 
  # filter students from the selected course/pres and finished the course
  filter(code_module == module, 
         code_presentation == presentation, 
         final_result != "Withdrawn") %>% 
  # drop the course identifiers ~ we focus on one course only
  select(-code_module,
         -code_presentation) 
```

The assessments needs to be reduced to the Tutor Marked Assigments only, because 
they represent milestones in the course:
```{r}
# list of assessments
selected_assessments <-
  assessment %>% 
  # filter assessments from the selected course/pres and focus on TMAs
  filter(code_module == module, 
         code_presentation == presentation,
         assessment_type == "TMA") %>% 
  # drop course identifiers
  select(-code_module,
         -code_presentation) %>% 
  # transform cutoff day to the week ~ we focus on data on level of weeks
  mutate(week = floor(date/7))  %>% 
  # shift the assessment ids to something more meaningful 
  mutate(id_assessment = id_assessment - 34872)

# student assessment results
selected_student_assessments <-
  student_assessment %>% 
  # shift the assessment ids to something more meaningful 
  mutate(id_assessment = id_assessment - 34872) %>% 
  # "filter the assessment to only assessments from selected course/pres
  semi_join(selected_assessments, by = "id_assessment") %>% 
  # select only relevant columns
  select(id_assessment,id_student,score)

```

Virtual learning environment data needs to be filtered and transformed too:
```{r}
# list of relevant vle resources
selected_vles <-
  vle %>% 
  # filter vle resources from the selected course/pres
  filter(code_module == module, code_presentation == presentation) %>%
  # remove irrelevant columns
  select(-week_from, -week_to)

# student vle interactions
selected_student_vles <-
  student_vle %>% 
  # filter the interactions to selected 
  inner_join(selected_vles) %>% 
  # transform day of interaction to week
  mutate(week = floor(date/7)) %>%
  # select relevant columns
  select(id_student, id_site, activity_type,week, sum_click) %>% 
  group_by(id_student,id_site,activity_type,week) %>% 
  # create interaction summary over the week 
  summarise(sum_click = sum(sum_click)) %>% 
  ungroup %>% 
  # correct the date
  mutate(week = week + 3) 
```

# Identification of top 5 VLE activities
The identification is based on sum of number of clicks for the activity across 
all students and weeks:
```{r}
# compute the total number of clicks
total_clicks <-
  selected_student_vles %>% 
  summarise(sum(sum_click)) %>% 
  extract2(1)

# select top 5 activity types
top_5_activity_types <-
  selected_student_vles %>% 
  group_by(activity_type) %>% 
  # compute proportion of clicks in activity type
  summarise(prop = sum(sum_click)/total_clicks) %>% 
  # order descending
  arrange(desc(prop)) %>% 
  # select 5
  top_n(5) %>% 
  # extract them to the numeric vector
  extract2(1) 
```

# Data merging
We focus on the last TMA and it is considered as dependent variable. Then we need
to select the VLE activity data before the TMA and merge all tibbles together. Finally,
some transformations are required: weighted sum of TMA 1-4, summary of `activity_type` clicks
in the VLE. Students, who did not submitted all assessments, were filtered.
```{r}
N <- 5

# find cutoff week
w <- 
  selected_assessments %>% 
  filter(id_assessment == N) %>% 
  extract2("week") + 3

# filter VLEs to the data from weeks before the assessment and to selected activity types
filtered_vle <- 
  selected_student_vles %>% 
  filter(week <= w, activity_type %in% top_5_activity_types) 
  
final_data <-
  selected_students %>% 
  # connect students with their results
  left_join(selected_student_assessments, by = "id_student") %>%
  # split results into 5 columns
  pivot_wider(names_from = id_assessment, 
              names_prefix = "assessment_", 
              values_from = score) %>% 
  # remove column caused by missing results of students
  select(-assessment_NA) %>%
  # prepare variable indicating that student did not submit all assignments
  rowwise() %>% 
  mutate(filt = anyNA(c_across(starts_with("assessment")))) %>%
  # filter students who did not submit all assignments
  filter(filt == FALSE) %>%
  select(-filt) %>% 
  left_join(filtered_vle, by = "id_student") %>% 
  # remove id_site, which is irrelevant for our study
  select(-id_site) %>% 
  # summarize the result of first 4 assignments
  mutate(cummulative_assessment_results = 
           0.125*assessment_1 + 
           0.125*assessment_2 + 
           0.25*assessment_3 + 
           0.25*assessment_4) %>% 
  select(-assessment_1,-assessment_2,-assessment_3,-assessment_4) %>% 
  # rename dependent variable
  rename(assessment_score = assessment_5) %>% 
  # summarize the VLE activity for each student and activity_type
  group_by(id_student, activity_type) %>% 
  mutate(sum_click = sum(sum_click)) %>% 
  ungroup() %>%
  # final filtering
  filter(week < 24) %>% 
  select(-week) %>% 
  distinct() %>% 
  # correction of typo in the data
  mutate(imd_band = if_else(imd_band == "10-20", "10-20%", imd_band)) %>% 
  # split clicks into the 5 columns regarding the activity_types
  pivot_wider(names_from = activity_type, 
              names_prefix = "clicks_", 
              values_from = sum_click, 
              values_fill = 0) %>% 
  # move dependent variable to the end of dataset
  relocate(assessment_score, .after = last_col())
```

# Store the result
```{r, eval=FALSE}
write_csv(final_data, paste0(module, "_", presentation, "_data.csv"))
```

